Denotational Semantics of Gradual Typing using Synthetic Guarded Domain Theory

### 2024-10-24 paper-new/discussion.tex
fix broken citations

This work has focused on the ``Natural'' semantics of casts , which
validate the full call-by-value $\eta$ laws for types, but other cast
semantics have been proposed such as eager
(\cite{herman-tomb-flanagan-2010}) and transient (\cite{transient})
cast semantics which trade off certain $\eta$ equalities for other
benefits such as early error detection or reduced runtime
overhead\cite{deepandshallowtypes,newlicataahmed19}. It would be a
good test of the generality of our semantic framework to see if it can
model these alternative semantics by adjusting the semantics of types
and casts, and providing proofs of weakened $\eta$ principles.

>>>>

This work has focused on the ``Natural'' semantics of casts , which
validate the full call-by-value $\eta$ laws for types, but other cast
semantics have been proposed such as eager
(\cite{herman-tomb-flanagan-2010}) and transient (\cite{transient})
cast semantics which trade off certain $\eta$ equalities for other
benefits such as early error detection or reduced runtime
overhead\cite{deepandshallowtypes,new-licata-ahmed2019}. It would be a
good test of the generality of our semantic framework to see if it can
model these alternative semantics by adjusting the semantics of types
and casts, and providing proofs of weakened $\eta$ principles.

### 2024-10-24 paper-new/intro.tex
fix broken citations

Providing a semantics for gradual typing is inherently complicated in
that it involves: (1) recursion and recursive types through the
presence of dynamic types, (2) effects in the form of divergence and
errors (3) relational models in capturing the graduality
property. Recursion and recursive types must be handled using some
flavor of domain theory. Effects can be modeled using monads in the
style of Moggi, or adjunctions in the style of
Levy\cite{moggi,levy}. Relational properties and their verification
lead naturally to the use of reflexive graph categories or double
categories\cite{reflgraphcats,doublecats}.

>>>>

Providing a semantics for gradual typing is inherently complicated in
that it involves: (1) recursion and recursive types through the
presence of dynamic types, (2) effects in the form of divergence and
errors (3) relational models in capturing the graduality
property. Recursion and recursive types must be handled using some
flavor of domain theory. Effects can be modeled using monads in the
style of Moggi, or adjunctions in the style of
Levy\cite{MOGGI199155,39155,levy99,levystacks}. Relational properties and their verification
lead naturally to the use of reflexive graph categories or double
categories\cite{dunphythesis,new-licata18}.
